---
title: "Finches"
author: "Mikako Inaba"
date: "10/9/2019"
output: pdf_document
---
  

```{r}
knitr::opts_chunk$set(echo = TRUE)
require(Sleuth3)
require(ggplot2)
require(tigerstats)
require(rstan)
library(Sleuth3)

finches <- case0201
```



## Exploratory analysis

The data consists of measurements of the beak depths of 178 finches, some measured before the draught, and some measured after the draught.

This is a bit of an overkill, but let's see from what years the data is:

```{r}
qplot(finches$Year, bins=3)
```

Now, let's see how the beak depths differ across the years
```{r message=FALSE}
ggplot(finches, aes(x=Depth)) + 
  geom_histogram(data=subset(finches, Year == 1976), fill = "red", alpha = 0.2) + 
  geom_histogram(data=subset(finches, Year == 1978), fill = "blue", alpha = 0.2)
```

Box plots and density plots can also be instructive
```{r boxdensityplots}
bwplot(Year ~ Depth, data = finches)
densityplot(~Depth, groups = Year, auto.key = TRUE, data = finches)
```

<!-------------------------------------------------------------------------->
  
  ## Difference of Means: t-test
  
  
```{r}
y1976 <- subset(finches, Year == 1976)$Depth
y1978 <- subset(finches, Year == 1978)$Depth
t.test(y1976, y1978, var.equal = F )
```


```{r bayes}
finch_model_stan <- "
data{
int<lower=0> N1; //number of points in example 1
int<lower=0> N2; //number of points in example 2
real sample1[N1]; 
real sample2[N2]; 
}

parameters{
real mu1;
real mu2;
real<lower=0> sigma1;
real<lower=0> sigma2;
}

model{
sample1 ~ normal(mu1, sigma1);
sample2 ~ normal(mu2, sigma2);
}"

adaptSteps = 1000            # Number of steps to "tune" the samplers.
burnInSteps = 5000           # Number of steps to "burn-in" the samplers.
nChains = 3                  # Number of chains to run.
numSavedSteps=12000          # Total number of steps in chains to save.
thinSteps=10                 # Number of steps to "thin" (1=keep every step).

finch_model <- stan_model(model_code = finch_model_stan, model_name = "finch_model")
finch_fit <- sampling(object=finch_model,
                      data = list(sample1=y1976, sample2=y1978,N1=length(y1976),N2=length(y1978)),
                      chains = nChains ,
                      iter = ( ceiling(numSavedSteps/nChains)*thinSteps
                               +burnInSteps ) , 
                      warmup = burnInSteps , 
                      thin = thinSteps ,
                      init = "random" ) 

samples <- extract(finch_fit)
```

Let's look at the posterior distribution for `mu1`

```{r}
hist(samples$mu1)
```

(Are the variances equal?)

```{r}
hist(samples$sigma1-samples$sigma2)
```

So what about the means?

```{r}
hist(samples$mu1-samples$mu2)
mean(samples$mu1-samples$mu2 >= 0)
```